---
title: "Accessing the e-Stat API"
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: setup
#| message: false
#| warning: false
library(jpstat)
library(tidyverse)
```

jpstat uses the e-Stat API's metadata retrieval (`getMetaInfo`) and
statistical data retrieval (`getStatsData`) functions to let you
download statistical tables.

Each statistical table on e-Stat is identified by a statsDataId. You
can find the statsDataId in the URL of the dataset information page
or the table/graph display page, such as this [example
page](https://www.e-stat.go.jp/en/dbview?sid=0000010101).

Here, we use the [System of Social and Demographic
Statistics](https://www.e-stat.go.jp/en/dbview?sid=0000010101) linked
above to retrieve the male/female population of Tokyo and Osaka for
2010 and 2015.

Using the e-Stat API requires an API key; set it with
`Sys.setenv(ESTAT_API_KEY = )` before running the code below yourself.
The output shown on this page was captured from a real run, but is not
re-executed when this article is built, so it doesn't require an API
key at build time.

First, pass the statsDataId (or a URL containing it) to `estat()` to
retrieve the metadata (the attribute information for the statistical
data).

```r
Sys.setenv(ESTAT_API_KEY = "Your appId")

ssds <- estat(
  statsDataId = "https://www.e-stat.go.jp/en/dbview?sid=0000010101",
  lang = "E"
)
ssds
#> # ☐ tab:   Observation Value           [1] <code, name, level>
#> # ☐ cat01: A Population and Households [594] <code, name, level, unit>
#> # ☐ area:  AREA                        [48] <code, name, level>
#> # ☐ time:  SURVEY YEAR                 [51] <code, name, level>
#> #
#> # Please `activate()`.
```

This table has four columns (referred to below as keys): `tab`,
`cat01`, `area`, and `time`. Each key carries the following
information:

1.  The default column name (e.g. `tab`)
2.  The number of items (e.g. `[1]`)
3.  Attributes such as code and name (e.g. `<code, name, level>`)

From here, we'll change the column name, item count, and attributes
for each key one at a time. To edit a key's information, select it
with `activate()`.

For example, activate the `tab` key like so:

```r
ssds |>
  activate(tab)
#> # ☒ tab:   Observation Value           [1] <code, name, level>
#> # ☐ cat01: A Population and Households [594] <code, name, level, unit>
#> # ☐ area:  AREA                        [48] <code, name, level>
#> # ☐ time:  SURVEY YEAR                 [51] <code, name, level>
#> #
#> # A tibble: 1 × 4
#> # Stickers: .estat_rowid
#>   .estat_rowid code  name              level
#> *        <int> <chr> <chr>             <chr>
#> 1            1 00001 Observation value 1

# Or
ssds |>
  activate(1)
#> # ☒ tab:   Observation Value           [1] <code, name, level>
#> # ☐ cat01: A Population and Households [594] <code, name, level, unit>
#> # ☐ area:  AREA                        [48] <code, name, level>
#> # ☐ time:  SURVEY YEAR                 [51] <code, name, level>
#> #
#> # A tibble: 1 × 4
#> # Stickers: .estat_rowid
#>   .estat_rowid code  name              level
#> *        <int> <chr> <chr>             <chr>
#> 1            1 00001 Observation value 1
```

Once a key is active, its item information is displayed, and you can
narrow it down with `filter()` and `select()`. Here, `tab` has only a
single item, so we simply drop the column with `select()`.

```r
ssds <- ssds |>
  activate(tab) |>
  # Only one item, so drop the column entirely
  select()
```

Next, select "Total population (Male)" and "Total population
(Female)" from `cat01`. `rekey()` renames the key (here, `sex`),
which also becomes the column name used when the data is downloaded.
As above, filter the items and keep the `name` and `unit` columns.

```r
ssds <- ssds |>
  activate(cat01) |>
  rekey("sex") |>
  filter(str_detect(name, "Total population \\((Male|Female)\\)")) |>
  select(name, unit)
```

Do the same for `area` (Tokyo-to and Osaka-fu) and `time` (2010 and
2015).

```r
ssds <- ssds |>
  activate(area) |>
  rekey("pref") |>
  filter(name %in% c("Tokyo-to", "Osaka-fu")) |>
  select(code, name) |>

  activate(time) |>
  rekey("year") |>
  filter(name %in% c("2010", "2015")) |>
  select(name)
```

With these steps, the column names, item counts, and attributes have
all been updated:

```r
ssds
#> # ☐ tab:  Observation Value           [1] <>
#> # ☐ sex:  A Population and Households [2] <name, unit>
#> # ☐ pref: AREA                        [2] <code, name>
#> # ☒ year: SURVEY YEAR                 [2] <name>
#> #
#> # A tibble: 2 × 2
#> # Stickers: .estat_rowid
#>   name  .estat_rowid
#> * <chr>        <int>
#> 1 2010            36
#> 2 2015            41
```

Finally, download the data with `collect()`. The `n` argument of
`collect()` sets the name of the value column.

```r
population <- ssds |>
  collect(n = "population")
#> The total number of data is 8.

knitr::kable(population)
```

| sex_name | sex_unit | pref_code | pref_name | year_name | population |
|:---|:---|:---|:---|:---|:---|
| A110101_Total population (Male) | person | 13000 | Tokyo-to | 2010 | 6512110 |
| A110101_Total population (Male) | person | 13000 | Tokyo-to | 2015 | 6666690 |
| A110101_Total population (Male) | person | 27000 | Osaka-fu | 2010 | 4285566 |
| A110101_Total population (Male) | person | 27000 | Osaka-fu | 2015 | 4256049 |
| A110102_Total population (Female) | person | 13000 | Tokyo-to | 2010 | 6647278 |
| A110102_Total population (Female) | person | 13000 | Tokyo-to | 2015 | 6848581 |
| A110102_Total population (Female) | person | 27000 | Osaka-fu | 2010 | 4579679 |
| A110102_Total population (Female) | person | 27000 | Osaka-fu | 2015 | 4583420 |
